Week 5

### 2024-10-02 Webscraping-Chapter8-tutorial.Rmd
indexrenewal

goal: get to know \
(i) who the Radboud University Department of Sociology staff is, \
(ii) what they publish with respect to scientific work, and \
(iii) who they collaborate with.

@@

# get anchor data - Leiden University - PolSci 
```{r}

@@

# sort rows into names and expertise 
```{r}
#Assigning even and odd rows to 'fodd' and 'feven' (bc the output just lists the left and right side of the table interchangably)
fodd <- function(x) x%%2 != 0
feven <- function(x) x%%2 == 0

@@

# Extract names and last names 
```{r, name extraction}
# gsub   is a function that remove something and replaces it with something else

@@

## clean up names 
```{r, clean up names }

@@

# Google Scholar Profiles and Publication
## reminder: for loops 
```{r, for loop}
# The 'for loop': for every i in a vector (can be numbers, strings, etc.), say 1 to 10, you can do
# 'something'
for (i in 1:10) {
    print(i)  # So for every i from 1 to 10, we print i, see what happens!
}

# # or do something more complicated p <- rnorm(10, 0, 1) # draw 10 normally distributed numbers
# with mean 0 and SD 1 (so z-scores, essentially) plot(density(p)) # relatively, normal, right?  u
# <- 0 # make an element we can fill up in the loop below for (i in 1:10) { u[i] <- p[i]*p[i] # get
# p-squared for every i-th element in vector p print(u[i]) # and print that squared element }
```
## linking a SINGLE staff member to their google scholar link 
```{r}

@@

```
### get scholar publications, citations
```{r, pubs and citations}
get_publications("Iu23-90AAAAJ")  # Jochem's pubs

get_citation_history("Iu23-90AAAAJ")  # Jochem's citation history
```
### get collaborators 
```{r, collaborators}
jochem_coauthors <- get_coauthors("Iu23-90AAAAJ", n_coauthors = 50, n_deep = 1)  
# Jochem's collaborators and their co-authors!

@@

## plot network of single staff member 
```{r, plot}
plot_coauthors(get_coauthors("Iu23-90AAAAJ", n_coauthors = 20, n_deep = 1), size_labels = 2)  
# Doesn't look like much yet, but we'll make it prettier later.

```
# link scholar profiles of ALL STAFF 
```{r, staff total scholar link, eval=FALSE, include=FALSE}
# Look throught get_scholar_id_fix(last_name, first_name, affiliation) 
# if we can find google scholar profiles of sociology staff!
for (i in 1:nrow(soc_df)) {

@@

# remove those without pubs from the df
# seems we're left with about 34 sociology staff members!
soc_df <- soc_df[!soc_df$gs_id == "", ]
soc_df
```
## get publications (all staff)
```{r eval=FALSE, include=FALSE}
#CURRENTLY: CODE IN THIS CHUNK IS NOT RUN (change in little setting weel at the top rightof this chunk)

@@

```
## google scholar says NO :(  (too many requests despite random requests)
-> cannot make lists with publications and profiles as of now (evening of 17th September) :/

# continue shortly before "getting collaborators" at Chapter 8.6.4
search and find: note how soc_list_profiles

# gender from names 
```{r}
# voornamenbank Maartens insitute 
# e.g. Vero:
# https://nvb.meertens.knaw.nl/naam/is/Vero

@@

# custom functions chunk for functions to run later on 
```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

@@

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```
# Open alex 
```{r}
# SEE SUNBELT WEBSITE: Attributes tab 
# if you put html from open alex api into the read html function then you need to replace any spaces with a + symbol

@@

```
# R Package openalexR (using r wrapper )
```{r}
# make JSON object easier to use

@@

```
## Adjacency matrix for 4 authors 
```{r}
packages <- c("tidyverse", "scholar", "openalexR", "rvest", "jsonlite")
fpackage.check(packages)
# make adjacency matrix for these four people

@@

```
### works

@@

```
## make matrix 
```{r}
# ALT + MINUS make the ARROWS <- 
# STR + SHIFT + M makes the %>% 
# STR + SHIFT + C can make a comment of multiple lines you highlighted

>>>>

goal: get to know\
(i) who the Radboud University Department of Sociology staff is,\
(ii) what they publish with respect to scientific work, and\
(iii) who they collaborate with.

@@

# get anchor data - Leiden University - PolSci

@@

# sort rows into names and expertise

```{r}
#Assigning even and odd rows to 'fodd' and 'feven' (bc the output just lists the left and right side of the table interchangably)
fodd <- function(x) x%%2 != 0
feven <- function(x) x%%2 == 0

@@

# Extract names and last names

```{r, name extraction}
# gsub   is a function that remove something and replaces it with something else

@@

## clean up names

```{r, clean up names }

@@

# Google Scholar Profiles and Publication

## reminder: for loops

```{r, for loop}
# The 'for loop': for every i in a vector (can be numbers, strings, etc.), say 1 to 10, you can do
# 'something'
for (i in 1:10) {
    print(i)  # So for every i from 1 to 10, we print i, see what happens!
}

# # or do something more complicated p <- rnorm(10, 0, 1) # draw 10 normally distributed numbers
# with mean 0 and SD 1 (so z-scores, essentially) plot(density(p)) # relatively, normal, right?  u
# <- 0 # make an element we can fill up in the loop below for (i in 1:10) { u[i] <- p[i]*p[i] # get
# p-squared for every i-th element in vector p print(u[i]) # and print that squared element }
```

## linking a SINGLE staff member to their google scholar link

@@

@@

### get scholar publications, citations

```{r, pubs and citations}
get_publications("Iu23-90AAAAJ")  # Jochem's pubs

get_citation_history("Iu23-90AAAAJ")  # Jochem's citation history
```

### get collaborators

```{r, collaborators}
jochem_coauthors <- get_coauthors("Iu23-90AAAAJ", n_coauthors = 50, n_deep = 1)  
# Jochem's collaborators and their co-authors!

## plot network of single staff member

```{r, plot}
plot_coauthors(get_coauthors("Iu23-90AAAAJ", n_coauthors = 20, n_deep = 1), size_labels = 2)  
# Doesn't look like much yet, but we'll make it prettier later.

# link scholar profiles of ALL STAFF

```{r, staff total scholar link, eval=FALSE, include=FALSE}
# Look throught get_scholar_id_fix(last_name, first_name, affiliation) 
# if we can find google scholar profiles of sociology staff!
for (i in 1:nrow(soc_df)) {

@@

# remove those without pubs from the df
# seems we're left with about 34 sociology staff members!
soc_df <- soc_df[!soc_df$gs_id == "", ]
soc_df
```

## get publications (all staff)

```{r eval=FALSE, include=FALSE}
#CURRENTLY: CODE IN THIS CHUNK IS NOT RUN (change in little setting weel at the top rightof this chunk)

@@

## google scholar says NO :( (too many requests despite random requests)

-\> cannot make lists with publications and profiles as of now (evening of 17th September) :/

# continue shortly before "getting collaborators" at Chapter 8.6.4

search and find: note how soc_list_profiles

# gender from names

```{r}
# voornamenbank Maartens insitute 
# e.g. Vero:
# https://nvb.meertens.knaw.nl/naam/is/Vero

@@

# custom functions chunk for functions to run later on

```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

@@

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}
```

# Open alex

```{r}
# SEE SUNBELT WEBSITE: Attributes tab 
# if you put html from open alex api into the read html function then you need to replace any spaces with a + symbol

@@

# R Package openalexR (using r wrapper )

```{r}
# make JSON object easier to use

@@

## Adjacency matrix for 4 authors

```{r}
packages <- c("tidyverse", "scholar", "openalexR", "rvest", "jsonlite")
fpackage.check(packages)
# make adjacency matrix for these four people

@@

### works

@@

## make matrix

```{r}
# ALT + MINUS make the ARROWS <- 
# STR + SHIFT + M makes the %>% 
# STR + SHIFT + C can make a comment of multiple lines you highlighted

### 2024-10-02 index.Rmd
updatedmenusweek2
indexrenewal

---
title: "Lab journal"
author: "by: Vero"
bibliography: references.bib

---

@@

----

<!--- The ---- is a visual line (page break looking mf) on the 
website 
Always ADD A SPACE BETWEEN A PAE BREAK AND the next comment, otherwise it will appear on page --->

<!--- This is the best way to write a comment. 
It also demonstrates, you can add html directly in your rmarkdown file. See: https://www.w3schools.com/html/html_comments.asp --->

<!--- to use a multi column page we use Pandoc's fenced Div. 
see here: https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html 
But please note, we are not here to make a beautiful website, our intention is just to have a quick lab journal up and running --->

:::: {.columns}
::: {.column width="50%"}

# My Lab Journal
Last compiled on `r format(Sys.time(), '%B, %Y')`

*Describe lab journal here.*

@@

----

@@

treeeeees

:::

::: {.column width="50%"}

![](images/cover.jpg)

:::
::::

## How to use

1. Fork the repository for this website and follow instructions on `README` to get started [here](https://github.com/JochemTolsma/labjournal).

2. Journal your work, by editing the .Rmd-files

3. ...

4. Change everything to make it your own!

Use the top menu to navigate through the journal.

----

# References

>>>>

---
title: "Lab journal"
author: "by: Verooo"
bibliography: references.bib
---

@@

```{r token, include=FALSE}
#install.packages("gitcreds")
# gitcreds::gitcreds_set()
```

# First page of something totally awesome yes.

[CTRL + ALT + I to make new chunks]{.smallcaps}

[ALT + MINUS makes \<-]{.smallcaps}

[STR + SHIFT + M makes the %\>%]{.smallcaps}

[STR + SHIFT + C can make a comment of multiple lines you highlighted]{.smallcaps}

------------------------------------------------------------------------

```{=html}
<!--- The ---- is a visual line (page break looking mf) on the 
website 
Always ADD A SPACE BETWEEN A PAGE BREAK AND the next comment, otherwise it will appear on page --->
```
```{=html}
<!--- This is the best way to write a comment. 
It also demonstrates, you can add html directly in your rmarkdown file. See: https://www.w3schools.com/html/html_comments.asp --->
```
```{=html}
<!--- to use a multi column page we use Pandoc's fenced Div. 
see here: https://bookdown.org/yihui/rmarkdown-cookbook/multi-column.html 
But please note, we are not here to make a beautiful website, our intention is just to have a quick lab journal up and running --->
```
::: columns
::: {.column width="50%"}
# Lab Journal

Last compiled on `r format(Sys.time(), '%B, %Y')`

*Description: Very cool lab journal. Yes, yes.*

@@

------------------------------------------------------------------------

@@

treeeeees
:::

::: {.column width="50%"}
![](images/krill-issue.jpg)
:::
:::

### 2024-10-02 networkvis.Rmd
w4heading
indexrenewal

---
title: "Webscraping my data 1"
author: "Verooo"
date: "2024-09-20"
output: html_document
---

---
title: "RQ draft 3"
author: "Verooo"
date: "2024-09-20"
output: html_document
---

@@

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

>>>>

---
title: "Chapter 9 - Network Visualisation Practice"
author: "Verooo"
date: "2024-09-25"
output: html_document
---

@@

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 9.1 Zachary’s karate club

```{r }
require(igraph)
g <- make_graph("Zachary")
plot(g)

# adjacentcy matrix 
gmat <- as_adjacency_matrix(g, type = "both", sparse = FALSE)
gmat
```

## 9.1.1. Descriptive statistics

### Size

```{r}
# number of nodes
vcount(g)
# number of edges
ecount(g)
```

#### degree

In-degree = how many ties come “towards you”

```{r}
igraph::degree(g)

hist(table(degree(g)), xlab='indegree', main= 'Histogram of indegree')
```

#### transitivity

For undirected networks, the clustering coefficient is the same as the **transitivity index:\
**the number of transitive triads divided by all possible transitive triads.\
For directed graphs not so.

```{r}
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero"))

see also: 9.1.3 triad-census

#### betweenness

For all dyads: testing shortest path.

Number of shortest path between two nodes\
divided by\
How many of those shortest paths is "i" part of ?

```{r}
igraph::betweenness(g, directed = FALSE)
```

## 9.1.2 dyad-census

```{r}
igraph::dyad_census(g)
```

##9.1.3 triad-census

```{r}
igraph::triad_census(g)
# I will use sna because it shows the names of the triads as well.
sna::triad.census(gmat)

unloadNamespace("sna")  #I will detach this package again, otherwise it will interfere with all kind of functions from igraph, and my students will hate me for that.

# calculate transitivity 
igraph::transitivity(g, type = "global")
sna::gtrans(gmat)

triad_g <- data.frame(sna::triad.census(gmat))

transitivity_g <- (3 * triad_g$X300)/(triad_g$X201 + 3 * triad_g$X300)
transitivity_g
# 0.2556818

## 9.1.4 Network visualisation

### proportional betweenness score

```{r}
# changing V
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 10  #after some trial and error
plot(g, mode = "undirected")
```

placing node 33 and 34 further apart:

```{r}
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)
```

place 33 and 34 even further apart:

```{r}
l  #let us take a look at the coordinates
l[1, 1] <- 4
l[34, 1] <- -3.5
plot(g, layout = l)
```

***CAUTION!!***

"In the last step we manipulated our data. The coordinates of two nodes have been altered manually and are no longer determined by some ‘objective’ algorithm. Is this not the same as slightly changing an inconvenient correlation in a correlation matrix or as removing some outliers from our dataset? At the very minimum, make sure to tell your readers what you have done! Add a note/legend to your graph."

```{r}
plot(g, layout = l, margin = c(0, 0, 0, 0))
legend(x = -2.5, y = -1.5, c("Note: the position of nodes 1 and 34 have been set by Jochem Tolsma \n for visualisation purposes only and do not reflect network properties"), bty = "n", cex = 0.8, xpd=TRUE)

PROBLEM: Note is not visible in my output

SOLUTION: added the "xpd = TRUE" to legend

## 9.2 Twittersphere in the Dutch HoP

```{r}
load("twitter_20190919.RData")  #change to your working directory
str(twitter_20190919, 1)
keyf <- twitter_20190919[[1]]
mydata <- twitter_20190919[[2]]
seats <- twitter_20190919[[3]]
```

So, what do we have?

-   `keyf`: a data.frame on 147 Dutch MPs.

-   `mydata`: This an object which is ready to analyze in RSiena. It is actually a quite complicated object. For now three things are important:

    1.  The nodes in `mydata` are the same as in `keyf` and in `seats`.

    2.  It contains the twitter data at three timepoints (in `mydata$depvars`). We have three layers:

        -   fnet: who follows whom

        -   atmnet: who atmentions whom

        -   rtnet: who retweats whom

    3.  It also contains time invariant information on the nodes (in `mydata$cCovars`)

-   `seats`: a dataset which contains the coordinates of the seats in the House of Parliament in the Netherlands.

### 2024-10-02 week5.Rmd
indexrenewal

---
title: "Chapter 9 - Network Visualisation Practice"
author: "Verooo"
date: "2024-09-25"
output: html_document
---

@@

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

@@

# 9.1 Zachary’s karate club

```{r }
require(igraph)
g <- make_graph("Zachary")
plot(g)

# adjacentcy matrix 
gmat <- as_adjacency_matrix(g, type = "both", sparse = FALSE)
gmat
```

## 9.1.1. Descriptive statistics

### Size

```{r}
# number of nodes
vcount(g)
# number of edges
ecount(g)
```

#### degree

In-degree = how many ties come “towards you”

```{r}
igraph::degree(g)

hist(table(degree(g)), xlab='indegree', main= 'Histogram of indegree')
```

#### transitivity

For undirected networks, the clustering coefficient is the same as the **transitivity index:\
**the number of transitive triads divided by all possible transitive triads.\
For directed graphs not so.

```{r}
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero"))

@@

see also: 9.1.3 triad-census

#### betweenness

For all dyads: testing shortest path.

Number of shortest path between two nodes\
divided by\
How many of those shortest paths is "i" part of ?

```{r}
igraph::betweenness(g, directed = FALSE)
```

## 9.1.2 dyad-census

```{r}
igraph::dyad_census(g)
```

##9.1.3 triad-census

```{r}
igraph::triad_census(g)
# I will use sna because it shows the names of the triads as well.
sna::triad.census(gmat)

unloadNamespace("sna")  #I will detach this package again, otherwise it will interfere with all kind of functions from igraph, and my students will hate me for that.

# calculate transitivity 
igraph::transitivity(g, type = "global")
sna::gtrans(gmat)

triad_g <- data.frame(sna::triad.census(gmat))

transitivity_g <- (3 * triad_g$X300)/(triad_g$X201 + 3 * triad_g$X300)
transitivity_g
# 0.2556818

## 9.1.4 Network visualisation

### proportional betweenness score

```{r}
# changing V
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 10  #after some trial and error
plot(g, mode = "undirected")
```

placing node 33 and 34 further apart:

```{r}
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)
```

place 33 and 34 even further apart:

```{r}
l  #let us take a look at the coordinates
l[1, 1] <- 4
l[34, 1] <- -3.5
plot(g, layout = l)
```

***CAUTION!!***

"In the last step we manipulated our data. The coordinates of two nodes have been altered manually and are no longer determined by some ‘objective’ algorithm. Is this not the same as slightly changing an inconvenient correlation in a correlation matrix or as removing some outliers from our dataset? At the very minimum, make sure to tell your readers what you have done! Add a note/legend to your graph."

```{r}
plot(g, layout = l, margin = c(0, 0, 0, 0))
legend(x = -2.5, y = -1.5, c("Note: the position of nodes 1 and 34 have been set by Jochem Tolsma \n for visualisation purposes only and do not reflect network properties"), bty = "n", cex = 0.8, xpd=TRUE)

PROBLEM: Note is not visible in my output

SOLUTION: added the "xpd = TRUE" to legend

## 9.2 Twittersphere in the Dutch HoP

```{r}
load("twitter_20190919.RData")  #change to your working directory
str(twitter_20190919, 1)
keyf <- twitter_20190919[[1]]
mydata <- twitter_20190919[[2]]
seats <- twitter_20190919[[3]]
```

So, what do we have?

-   `keyf`: a data.frame on 147 Dutch MPs.

-   `mydata`: This an object which is ready to analyze in RSiena. It is actually a quite complicated object. For now three things are important:

    1.  The nodes in `mydata` are the same as in `keyf` and in `seats`.

    2.  It contains the twitter data at three timepoints (in `mydata$depvars`). We have three layers:

        -   fnet: who follows whom

        -   atmnet: who atmentions whom

        -   rtnet: who retweats whom

    3.  It also contains time invariant information on the nodes (in `mydata$cCovars`)

-   `seats`: a dataset which contains the coordinates of the seats in the House of Parliament in the Netherlands.

>>>>

---
title: "Week 5"
author: "Verooo"
date: "2024-10-02"
output: html_document
---

@@

```{r klippy, echo=FALSE, include=FALSE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

@@

# Notes while discussing other peeps plots
Using RSiena for evolutin of network 
Wave, Ego alters 
RSciena: Ego, alter, wave 
Reorganize into an array for RSciena

Set size of isolates to SMALL 
put all isolates in one place 
then you have enough room left to have a more "zoomed in" frame for the actual collaborator networks

Or just remove all isolates lol 
```{r}

@@
